\chapter{Acknowledgements}
Completing a PhD is no small task, and is certainly not performed in isolation. I am lucky to have been supported by a wide variety of people over the years, many of whom I will have inevitably forgotten. I apologise if you are unintentionally left out of these acknowledgements.

First, thanks go to all of my teachers over the past twenty-six years, but especially those at Alleyn's School: Ed Mann, Anne Poole, Neil Kinnear, Sara Hopley, Andy Skinnard \& Margaret Hunnaball, and those at Cambridge: Rafi Blumenfeld, Derek Barnes, Adam Thorne, David Summers, Ulrich Kaiser, Steve Gull, Paul Alexander \& John Ellis. Having now been a supervisor myself, I know how difficult a student I must have been, and with that in mind I am especially grateful.

Many thanks are due to the moderators and posters on the various technical fora throughout the internet, particularly on \mbox{\href{http://stackexchange.com/}{StackExchange}}. You are the unsung heroes and heroines of the modern internet age.

I am very grateful for the care of the Astrophysics administrator Karen Scrivener. Paul has often said that you are the most important member of the department, and I entirely agree.

To my office mates; Sonke Hee, Richard Wolstenhulme and Do Young Kim for the many amusing, interesting and fruitful conversations over the past few years. 

To Professors Anthony Lasenby and Mike Hobson. One could not wish for more supportive and caring supervisors. I can only hope that one day I will be able to pass the favour on to the next generation of students with half as much skill and grace.

To my mother and father for their continuing love and support, be it financial, emotional or moral.

Last and most to my fianc\'{e}e Sophie Lovick, for her patience and sacrifice, and above all for managing to live with a Physicist for five and a half years.
